\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

%All praise and thanks are due to Almighty God, whose guidance and blessings have enabled me to complete this thesis. Without His infinite mercy, none of this would have been possible. \\

%\noindent
All praise and thanks are due to Almighty God, whose guidance and blessings have made the completion of this thesis possible. Without His infinite mercy, none of this would have been achieved.  \\

%I would like to express my deepest gratitude to my parents, whose unwavering love, support, and sacrifices have made this achievement possible. Their continuous encouragement and belief in me have been an invaluable source of strength.

\noindent
I would like to express my deepest gratitude to my parents. Their unwavering love, support, and sacrifices have been instrumental in my success. Their continuous encouragement and unwavering belief in me have been an invaluable source of strength. \\

%To my family, thank you for your endless support, understanding, and patience. Even though you were far away in my home country throughout my study journey, you remained my greatest source of strength. Your love, prayers, and unwavering support from afar have been a constant motivation for me to persevere.

\noindent
To my family, thank you for your endless support, understanding, and patience. Although geographically distant during my studies, you remained my greatest source of strength. Your love, prayers, and unwavering support from afar have been a constant source of motivation.\\

\noindent
I am deeply indebted to my supervisors, Professor Ana Aguiar and Professor João Mendes-Moreira, for their invaluable guidance, patience, and expertise. Their insightful feedback, constructive criticism, and unwavering support have been instrumental in shaping this research. I am truly grateful for their dedication and the opportunities they have provided me. \\

%I would also like to express my sincere thanks to all the members of the IT-Lab at FEUP. Their collaborative spirit, helpful discussions, and friendly camaraderie have created a stimulating and supportive research environment.

\noindent
I would also like to express my sincere thanks to all members of the IT-Lab at FEUP. Their collaborative spirit, helpful discussions, and friendly camaraderie fostered a stimulating and supportive research environment. \\

%\noindent
%Finally, I would like to acknowledge the \textit{Instituto de Telecomunicações} for their financial support, which made this research possible. \\

\noindent
Finally, I express my sincere appreciation to \textit{Instituto de Telecomunicações (IT)} for providing me with excellent research facilities, a supportive environment, and the resources necessary for the successful completion of my work. \\

\noindent
This research would not have been feasible without the crucial support provided by the following projects: S2MovingCity (CMUP-ERI/TIC/0010/2014),   Safe Cities (POCI-01-0247-FEDER-041435),  and Route 25 (C645463824-00000063).


\vspace{40mm}
\flushleft{Akilu Rilwan Muhammad } %\\
   %       Porto, \today }


%%%%%%%%%%%%%%  Backup
%All praise and thanks are due to Almighty God, whose guidance and blessings have enabled me to complete this thesis. Without His infinite mercy, none of this would have been possible.

%I would like to express my deepest gratitude to my parents, Alhaji Rilwan Muhammad and Zulai Ahmad, whose unwavering love, support, and sacrifices have made this achievement possible. Their continuous encouragement and belief in me have been an invaluable source of strength.

%To my family, thank you for your endless support, understanding, and patience. Even though you were far away in my home country throughout my academic journey, you remained my greatest source of strength. Your love, prayers, and steadfast support from afar have been a constant motivation for me to persevere.

%I am deeply indebted to my supervisors, Professor Ana Aguiar and Professor João Mendes-Moreira, for their invaluable guidance, patience, and expertise. Their insightful feedback, constructive criticism, and unwavering support have been instrumental in shaping this research. I am truly grateful for their dedication and the opportunities they have provided me.

%I would also like to express my sincere thanks to all the members of the IT-Lab at FEUP. Their collaborative spirit, helpful discussions, and friendly camaraderie have created a stimulating and supportive research environment.

%Finally, I would like to acknowledge the \textit{Instituto de Telecomunicações} for their financial support, which made this research possible.